---
title: "ford-demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ford-demo}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(FORD)   
library(FOCI)
```

# Introduction

In this vignette, we demonstrate **FORD**, a forward stepwise variable selection algorithm based on the integrated $R^2$ dependence measure. **FORD** is designed for variable ranking in both linear and nonlinear multivariate regression settings.

**FORD** closely follows the structure of **FOCI** [*A Simple Measure Of Conditional Dependence*](https://arxiv.org/pdf/1910.12327.pdf), but replaces the core dependence measure with **irdc**.

---

# Algorithm

Let $Y$ be the response variable and $\mathbf{X} = (X_1, \dots, X_p)$ the predictor variables. Given $n$ i.i.d. samples of $(Y, \mathbf{X})$, FORD proceeds as follows:

1. Select $j_1 = \arg\max_j \nu_n(Y, X_j)$  
   If $\nu_n(Y, X_{j_1}) \leq 0$, return $\hat{V} = \emptyset$

2. Iteratively add the feature that gives the **maximum increase** in irdc:
   $$
   j_{k+1} = \arg\max_{j \notin \{j_1, \ldots, j_k\}} \nu_n(Y, (X_{j_1}, \ldots, X_{j_k}, X_j))
   $$

3. Stop when the irdc does not increase anymore:
   $$
   \nu_n(Y, (X_{j_1}, \ldots, X_{j_k}, X_{j_{k+1}})) \leq \nu_n(Y, (X_{j_1}, \ldots, X_{j_k}))
   $$

If no such $k$ exists, select all variables.

---

# Example 1 — Complex nonlinear function of first 4 features

Here, $Y$ depends only on the first 4 features of $X$ in a nonlinear way.

```{r example-1}
set.seed(42)
n <- 2000
p <- 100
X <- matrix(rnorm(n * p), ncol = p)
colnames(X) <- paste0("X", seq_len(p))
Y <- X[, 1] * X[, 2] + sin(X[, 1] * X[, 3]) + X[, 4]^2
```

## FOCI Result

```{r foci-result1}
result_foci_1 <- foci(Y, X, numCores = 1)
result_foci_1
```

## FORD Result

```{r ford-result1}
result_ford_1 <- ford(Y, X, numCores = 1)
result_ford_1
```

---

# Example 2 — Selecting a fixed number of variables

We can force both FOCI and FORD to select a specific number of variables instead of using an automatic stopping rule.

## FOCI with 5 selected features

```{r foci-result2}
result_foci_2 <- foci(Y, X, num_features = 5, stop = FALSE, numCores = 1)
result_foci_2
```

## FORD with 5 selected features

```{r ford-result2}
result_ford_2 <- ford(Y, X, num_features = 5, stop = FALSE, numCores = 1)
result_ford_2
```

---

# Conclusion

**FORD** provides an interpretable, irdc-based alternative to FOCI for variable selection in regression tasks. It offers a principled forward selection framework that can detect complex nonlinear relationships and be adapted for fixed-size feature subsets.

For further theoretical details, see our paper:  
Azadkia and Roudaki (2025), [*A new measure of dependence: Integrated $R^2$*](https://arxiv.org/pdf/1910.12327.pdf).
